Einstein's gravity from a polynomial affine model

### 2015-07-29 AffGR-PRD.tex
Askirz corrections, about compare with Maxwell

\hypersetup{%
  pdftitle={Einstein's Gravity from an Affine Model},
  pdfauthor={Oscar Castillo-Felisola,}{Aureliano Skirzewski},
  pdfkeywords={Affine Gravity,} {Torsion,} {Generalised Gravity.},
  pdflang={English}
}

@@

\title{Einstein's gravity from an affine model}

@@

%% --------- Abstract
\begin{abstract}
  We show that the effective field equations for a recently formulated affine model of gravity, in the sector of a metric-compatible (torsion-free) connection, accept general Einstein manifolds ---with or without cosmological constant--- as solutions. Moreover, these effective field equations coincide with those obtained from a gravitational Yang--Mills theory known as Stephenson--Kilmister--Yang theory. Additionally, we find an equivalence between a minimally coupled massless scalar field in General Relativity with a ``minimally'' coupled scalar field in this affine model.
\end{abstract}

\pacs{02.40.Ma,04.50.Kd,04.90.+e}
\keywords{Affine Gravity, Torsion, Generalised Gravity.}

@@

Classically, General Relativity is the most successful theory of gravitational interactions. However, it is believed to be the effective theory of a, yet to be discover, more fundamental one. Among the problems of General Relativity we mention a few aspects related with the quantization: First, the fact that in the standard formalism of quantization General Relativity is non-renormalizable~\cite{'tHooft:1973us,'tHooft:1974bx,Deser:1974cz,Deser:1974cy}. Second, the Wheeler--DeWitt equations are not well-defined in general due to their non-polynomial dependence on the Arnowitt--Deser--Misner (or ADM) variables~\cite{Arnowitt:1959ah,Arnowitt:1960es,WheelerGeo,DeWitt:1967yk,DeWitt:1967ub,DeWitt:1967uc}. One of the most successful quantization methods of gravitational interactions ---developed from ideas by Gambini and Trias~\cite{Gambini:1980yz,Gambini:1986ew}, and Ashtekar~\cite{Ashtekar:1986yd,Ashtekar:1987gu}---, hints that General Relativity is not as complete as expected, since the Lagrangian formulation includes (at least) an extra term known as Holst term~\cite{Holst:1995pc}. On the other hand, it is well known that in the standard formulation of General Relativity the skew-symmetric part of the connection (or torsion) is assumed to be zero, albeit it is known that such assumption can be relaxed once matter is added, in particular fermionic matter.

Standing of the argument of the necessity of a quantum theory of gravity, several generalizations of Einstein's General Relativity have been proposed. One important class of these generalizations consider the connection field to be independent of the metric, known as Palatini type theories. Within this class of theories one finds the generalization due to Cartan, which consider a connection with nonvanishing torsion~\cite{Cartan1922,Cartan1923,Cartan1924,Cartan1925}. Interestingly, gravitational theories with generic connections have degrees of freedom which do not correspond to the expected massless graviton~\cite{Sezgin:1979zf}. Albeit these gravitational theories with an arbitrary connection posses ghosts, there are examples of well-behaved (stable) systems which have ghosts, such are the cases of the Pais--Uhlenbeck oscillator~\cite{Mannheim:2004qz,Bender:2007wu,Smilga:2008pr,Ilhan:2013xe} or the higher derivative supersymmetric quantum mechanical system presented in Ref.~\cite{Robert:2008}.

Therefore, under the premises of non-completeness and non-polynomial structure of the theory, 
%%%%%%%%% THE MODEL %%%%%%%%%
in a previous paper~\cite{Skirzewski:2014eta}, we presented a polynomial (purely) affine gravitational model in four dimensions built up entirely on the basis of full diffeomorphism invariance, and restricting ourselves to consider only naive power-counting renormalizable terms (which by no means guaranty the renormalizability of the system). We showed that its non-relativistic limit around a homogeneous and isotropic spacetime yields a Newtonian gravitational potential, despite the existence of torsion. We also highlighted the possibility of using standard methods to quantize the model, due to its polynomial structure (unlike the earliest proposals~\cite{Eddington1923math,schrodinger1950space}), and also the likelihood of avoiding the uniqueness of diffeomorphism-invariant states from Loop Quantum Gravity programme~\cite{Lewandowski:2005jk}.

@@

First of all, we highlight that the model constructed below has as fundamental field an affine connection, and no metric fields is needed for neither contract nor lowering or raising indices. Moreover, in order to guarantee the correct transformation of the Lagrangian density, the geometrical objects used to write down the action will be the curvature and torsion of an affine connection, $\hat{\Gamma}^\mu{}_{\rho\sigma}$, which accepts a decomposition on irreducible components as

@@

where ${\Gamma}^\mu{}_{\rho\sigma}$ is symmetric in the lower indices, $A_\mu$ is a vector field corresponding to the trace of torsion, and  $T^{\mu,\lambda\kappa}$ is a Curtright field~\cite{Curtright:1980yk}, satisfying $T^{\kappa,\mu\nu } = - T^{\kappa,\nu\mu }$ and $\epsilon_{\lambda\kappa\mu\nu}T^{\kappa,\mu\nu }=0$. Since no metric is present  the epsilon symbols are not related by raising (lowering) their indices, but instead we demand that \mbox{$\epsilon^{\delta\eta\lambda\kappa}\epsilon_{\mu\nu\rho\sigma}=4!\delta^{\delta}{}_{[\mu}\delta^\eta{}_{\nu}\delta^{\lambda}{}_{\rho} \delta^\kappa{}_{\sigma]}$.}

>>>>

\hypersetup{%
  pdftitle={Einstein's gravity from a polynomial affine model},
  pdfauthor={Oscar Castillo-Felisola,}{Aureliano Skirzewski},
  pdfkeywords={Affine Gravity,} {Torsion,} {Generalised Gravity.},
  pdflang={English}
}

@@

\title{Einstein's gravity from a polynomial affine model}

@@

%% --------- Abstract
\begin{abstract}
  We show that the effective field equations for a recently formulated polynomial affine model of gravity, in the sector of a torsion-free connection, accept general Einstein manifolds ---with or without cosmological constant--- as solutions. Moreover, these effective field equations coincide with those obtained from a gravitational Yang--Mills theory known as Stephenson--Kilmister--Yang theory. Additionally, we find a generalization of a minimally coupled massless scalar field in General Relativity with a ``minimally'' coupled scalar field in this affine model.
\end{abstract}

\pacs{02.40.Ma,04.50.Kd,04.90.+e}
\keywords{Affine Gravity, Torsion, Generalized Gravity.}

@@

Classically, General Relativity is the most successful theory of gravitational interactions. However, it is believed to be the effective theory of a, yet to be discovered, more fundamental one. Among the problems of General Relativity we mention a few aspects related with the quantization: First, the fact that in the standard formalism of quantization General Relativity is non-renormalizable~\cite{'tHooft:1973us,'tHooft:1974bx,Deser:1974cz,Deser:1974cy}. Second, the Wheeler--DeWitt equations are not well-defined in general due to their non-polynomial dependence on the Arnowitt--Deser--Misner (or ADM) variables~\cite{Arnowitt:1959ah,Arnowitt:1960es,WheelerGeo,DeWitt:1967yk,DeWitt:1967ub,DeWitt:1967uc}. One of the most successful quantization methods of gravitational interactions ---developed from ideas by Gambini and Trias~\cite{Gambini:1980yz,Gambini:1986ew}, and Ashtekar~\cite{Ashtekar:1986yd,Ashtekar:1987gu}---, hints that General Relativity is not as complete as expected, since the Lagrangian formulation includes (at least) an extra term known as Holst term~\cite{Holst:1995pc}. On the other hand, it is well known that in the standard formulation of General Relativity the skew-symmetric part of the connection (or torsion) is assumed to be zero, albeit it is known that such assumption can be relaxed once matter is added, in particular fermionic matter.

Standing on the argument of the necessity of a quantum theory of gravity, several generalizations of Einstein's General Relativity have been proposed. One important class of these generalizations considers the connection field to be independent of the metric, known as Palatini type theories. Within this class of theories one finds the generalization due to Cartan, which considers a connection with nonvanishing torsion~\cite{Cartan1922,Cartan1923,Cartan1924,Cartan1925}. Interestingly, gravitational theories with generic linear connections have degrees of freedom which do not correspond to the expected massless graviton~\cite{Sezgin:1979zf}. Albeit these gravitational theories with an arbitrary connection possess ghosts, there are examples of well-behaved (stable) systems which have ghosts, such are the cases of the Pais--Uhlenbeck oscillator~\cite{Mannheim:2004qz,Bender:2007wu,Smilga:2008pr,Ilhan:2013xe} or the higher derivative supersymmetric quantum mechanical system presented in Ref.~\cite{Robert:2008}.

Therefore, under the premises of non-completeness and non-polynomial structure of the theory, 
%%%%%%%%% THE MODEL %%%%%%%%%
in a previous paper~\cite{Skirzewski:2014eta}, we presented a polynomial (purely) affine gravitational model in four dimensions built up entirely on the basis of full diffeomorphism invariance, and restricting ourselves to consider only naive power-counting renormalizable terms (which by no means guarantees the renormalizability of the system). We showed that its non-relativistic limit around a homogeneous and isotropic spacetime yields a Newtonian gravitational potential, despite the existence of torsion. We also highlighted the possibility of using standard methods to quantize the model, due to its polynomial structure (unlike the earliest proposals~\cite{Eddington1923math,schrodinger1950space}), and also the likelihood of avoiding the uniqueness of diffeomorphism-invariant states from Loop Quantum Gravity programme~\cite{Lewandowski:2005jk}, due to the absence of a fundamental metric.

@@

First of all, we highlight that the model constructed below has as fundamental field an affine connection, and no metric fields is needed for neither contracting nor lowering or raising indices. Moreover, in order to guarantee the correct transformation of the Lagrangian density, the geometrical objects used to write down the action will be the curvature and torsion of an affine connection, $\hat{\Gamma}^\mu{}_{\rho\sigma}$, which accepts a decomposition on irreducible components as

@@

where ${\Gamma}^\mu{}_{\rho\sigma}$ is symmetric in the lower indices, $A_\mu$ is a vector field corresponding to the trace of torsion, and  $T^{\mu,\lambda\kappa}$ is a Curtright field~\cite{Curtright:1980yk}, satisfying \mbox{$T^{\kappa,\mu\nu } = - T^{\kappa,\nu\mu }$} and $\epsilon_{\lambda\kappa\mu\nu}T^{\kappa,\mu\nu }=0$. Since no metric is present  the epsilon symbols are not related by raising (lowering) their indices, but instead we demand that \mbox{$\epsilon^{\delta\eta\lambda\kappa}\epsilon_{\mu\nu\rho\sigma}=4!\delta^{\delta}{}_{[\mu}\delta^\eta{}_{\nu}\delta^{\lambda}{}_{\rho} \delta^\kappa{}_{\sigma]}$.}
